---
title: "campaign finance"
output:
---

FEC bulk data URL https://www.fec.gov/
Data from the 2021 - 2022 election cycle

```{r}
library(tidyverse)
library(lubridate)
```

```{r}
transactions <- read_csv("data/transactions.csv")
transactions_types <- read_csv("data/transaction_types.csv")
candidates <- read_csv("data/candidates.csv")
committees <- read_csv("data/committees.csv")
```

What transactions types are we working with
```{r}
transactions %>% 
  count(transaction_tp)
```

Join to transaction_types
Test first
```{r}
transactions_types %>% 
  count (transaction_tp) %>% 
  filter(n>1)

#no duplicate
```

Use a left join
```{r}
# Use filter(is.na(transaction_tp)) to check nulls
transactions %>% 
  left_join(transactions_types, by = "transaction_tp") %>% 
  filter(is.na(transaction_tp))

# Check the code and desc matchups
transactions %>% 
  left_join(transactions_types, by = "transaction_tp") %>% 
  count(transaction_tp, transaction_desc)
#Exactly 10896 rows. Good sign, now check nulls on the left table 
#Count only gives you the number of rows. 

#Optional step is to create a new table from the join
transactions2 <- transactions %>% 
  left_join(transactions_types, by = "transaction_tp")
```

How much money was spent each transaction type
```{r}
transactions %>% 
  left_join(transactions_types, by = "transaction_tp") %>% 
  group_by(transaction_desc) %>% 
  summarise(total_amt = sum(transaction_amt), n=n()) %>% #n: how many times it appears
  arrange(desc(total_amt))

```

Which candidates are in our data
```{r}
transactions %>% 
  count(cand_id) %>% 
  arrange(desc(n)) %>% 
  left_join(candidates, by = "cand_id")
```
Which candidate had the most money spent on them in independent expenditures?
```{r}
transactions %>% 
  filter(transaction_tp == "24A" | transaction_tp == "24E") %>% 
  left_join(candidates, by = "cand_id") %>% 
  group_by(cand_name) %>% 
  summarise(total_amt = sum(transaction_amt)) %>% 
  arrange(desc(total_amt))
```
```{r}
#Look at opposing
transactions %>% 
  filter(transaction_tp == "24A") %>% 
  left_join(candidates, by = "cand_id") %>% 
  group_by(cand_name) %>% 
  summarise(total_amt = sum(transaction_amt)) %>% 
  arrange(desc(total_amt))

#Look at advocating
transactions %>% 
  filter(transaction_tp == "24E") %>% 
  left_join(candidates, by = "cand_id") %>% 
  group_by(cand_name) %>% 
  summarise(total_amt = sum(transaction_amt)) %>% 
  arrange(desc(total_amt))
```
Can we make it more efficient?
```{r}
transactions %>% 
  filter(transaction_tp %in% c("24A", "24E")) %>% 
  left_join(candidates, by = "cand_id") %>% 
  group_by(cand_name, transaction_tp) %>% 
  summarise(total_amt = sum(transaction_amt)) %>% 
  pivot_wider(names_from = transaction_tp, values_from = total_amt) %>% #Use it when you want your data longer or wider
  mutate(total = sum(`24A`,`24E`, na.rm = T), opposing_pct = `24A`/total) %>% 
  arrange(desc(opposing_pct))
```

Who are the committees spending this independt expenditure money?
```{r}
transactions %>% 
  filter(transaction_tp %in% c("24A", "24E")) %>% 
  left_join(committees, by="cmte_id") %>% 
  group_by(cmte_nm) %>% 
  summarise(total = sum(transaction_amt)) %>% 
  arrange(desc(total))
```

