---
title: "Columbia Dispatch"
output: html_notebook
---

Columbia Police Dispatch
https://www.como.gov/CMS/911dispatch/police.php
Downloaded 1/29/2024 8:43 a.m.
1/1/2019 to 1/29/2024

This is what we requested, but we actually only got
1/29/2023 to 1/29/2024

DO NOT DO AGAIN. LIBRARY INSTALLED
```{r}
# install.packages("lubridate")
```

```{r}
library(tidyverse)
library(lubridate)
```
Load the data
```{r}
dispatch<-read_csv("data/como_dispatch.csv")
```
83,190 rows
1 row = 1 call that resulted dispatch

Assessing the data
```{r}
dispatch %>% 
  group_by(CallDateTime, Address) %>% 
  summarise(n=n()) %>% 
  filter (n>1)

dispatch %>% 
  filter(CallDateTime == "10/8/2023 2:08:54 PM")

dispatch %>% 
  group_by(InNum) %>% 
  summarise(n = n()) %>% 
  filter(n>1)

dispatch %>% 
  count(InNum) %>% 
  arrange(desc(n))
  
dispatch %>% 
  filter(InNum=="2023029401")
```
There are exact duplicates! Call the city
Possible reason: Multiple officer were dispatched, system enter things twice

Let's say we learned that exact duplicates are mistakes and should be removed:
```{r}
dispatch <- dispatch %>% 
  distinct()
```

Assess every column:
```{r}
dispatch %>% 
  count(InNum) %>% 
  arrange(desc(n))
#Still some duplicates



```

Create a true date column for CallDateTime
```{r}
dispatch %>% 
  mutate(new_date = mdy_hms(CallDateTime)) %>% 
  select(CallDateTime, new_date)

dispatch <- dispatch %>% #Overwrite dispatch
  mutate(new_date = mdy_hms(CallDateTime)) 
```

What's the date range of our data
```{r}
range(dispatch$new_date)
#Use the $ to turn it into a function
#[1] "2023-01-29 08:47:18 UTC" "2024-01-29 02:29:40 UTC", however, we requested five-year data but only got 2023-2024
```

Other ways to work with true dates
```{r}
#how many incident happened each year/month?
dispatch %>% 
  count(month(new_date)) 

#month(new_date) Pull the month out of each values, then count() will count the month

#nesting:getting information from one to the next. Inword do the first, and next function do the next
```

