---
title: "R Notebook"
output: html_notebook
---

```{r}
# don't need to run this code again: 
install.packages("tidyverse").
```

We load libraries at the beginning of every script
```{r}
library(tidyverse)
```

Load some data
```{r}
weather <- read_csv("data/weather_events.csv")
```

Main tidyverse functions
```{r}
# SORTING
weather %>% arrange(desc(date))

weather %>% arrange(desc(injuries))

# FILTERING
weather %>% filter(type == "tornado") %>% arrange(desc(fatalities))

weather %>% filter(fatalities > 0)

# operators
# ==
# !=
# >, <, >=, <=
# AND, OR: &, |

weather %>% filter(type == "tornado" & fatalities > 0)

weather %>% filter(type == "tornado" | type == "wind")

# GROUPING
weather %>% 
  group_by(type) %>% 
  summarise(total_inj = sum(injuries)) %>% 
  arrange(desc(total_inj))

weather %>% 
  summarise(total_inj = sum(injuries)) %>% 
  arrange(desc(total_inj))

```

To answer these questions, you'll need the summary function that counts rows: n()
This function requires no arguments. ex: `summarise(number_of_events = n())`
Remember to show your work and answer the question (if you can). 

QUESTIONS FOR HOMEWORK: 
1. What was the most common weather event type in your home state?
```{r}
weather %>% 
  filter(state == "MO") %>% 
  group_by(type) %>% 
  summarise(event_num = n()) %>% 
  arrange(desc(event_num))
```
Hail is the most common weather in MO with a number of 17957.

2. How many injuries and fatalities were associated with it?
```{r}
weather %>% 
  filter(state == "MO" & type == "hail") %>% 
  summarise(total_inj = sum(injuries), total_fatal = sum(fatalities))
```
Total injuries is 7

3. Which weather event type has the highest fatality rate? 
```{r}
weather %>% 
  group_by(type) %>% 
  summarize(avg_fatal = mean(fatalities)) %>% 
  arrange(desc(avg_fatal))
```

4. Which county has the most weather events? (use `cnty1_fips`)
```{r}
weather %>% 
  group_by(state, cnty1_fips) %>% 
  summarise(cnty = n()) %>% 
  arrange(desc(cnty))
```
County 3 in KS has the most weather events.

5. Are weather events increasing over time? 
```{r}
weather %>% 
  group_by(year) %>% 
  summarise(number_of_events = n()) %>% 
  arrange(desc(number_of_events))
```

